\title{A Survey of Recent Work on Designing Competitive Online Algorithms using the Primal-Dual Method \\ \begin{large}6.854 Advanced Algorithms Project\end{large}}

\author{Fredrik Kjolstad \\ fredrikk@mit.edu \and Ludwig Schmidt \\ ludwigs@mit.edu \and Christos Tzamos \\ tzamos@mit.edu}
\date{}

\documentclass[10pt, twocolumn]{article}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{url}
\usepackage{relsize}
\usepackage{xspace}
\usepackage{subfigure}
\usepackage{graphicx,color}
\usepackage[noend]{algorithmic}
\usepackage{algorithm}
\usepackage{paralist}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}


\usepackage{fullpage}

% Todos
\newcommand{\fred}[1]{%
	\textcolor{red}{[#1]}
}
\newcommand{\christos}[1]{%
	\textcolor{red}{[#1]}
}
\newcommand{\ludwig}[1]{%
	\textcolor{red}{[#1]}
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}
\maketitle

\begin{abstract}
Online algorithms have become increasingly popular in the last few decades.
An important reason for this is that they capture the uncertainty we face in many important domains, ranging from computational finance to internet security and catastrophe management.
The primal-dual method is an approach that has been gaining popularity as a technique to arrive at approximations for NP-hard problems.
More recently it has also been applied as a general framework to solve many online algorithms.
In 2009, Buchbinder and Naor published a survey of applications of the primal-dual method to online algorithms.
Since then new complex online problems have been tackled using this technique, such as the online node-weighted steiner tree problem, the k-server problem and online job-migration.
We present a survey of recent applications of the primal-dual method to online problems.
\end{abstract}

\section{Introduction}
\label{intro}
\input{intro}



\section{The Primal-Dual Approach}
\label{primal-dual}

In this section we present the primal-dual approach.
First, we explain the required background in linear programming and duality.
We then introduce the technique using the well known ski rental problem as an example.
Finally, we discuss the general primal--dual framework so that it can be applied to other problems.

We assume the reader is already familiar with the online algorithm model, competitive analysis, approximation algorithms, (integer) linear programming, randomized rounding and the concepts of weak and strong duality and will not review these concepts here.
For a brief discussion of these concepts we refer the reader to section~2 of Buchbinder and Naor's survey~\cite{buchbinder09:survey}.

\input{preliminaries}
\input{ski-rental}


\section{\mbox{Graph Optimization}}
\label{steiner-tree}
\input{steiner-tree}

\section{Towards the Randomized k-server Conjecture: a Primal-Dual Approach}
\label{k-server}
\input{k-server}

\section{Job-Migration}
\label{job-migration}
\input{job-migration}

\section{Conclusions}
\fred{TODO}

\bibliographystyle{abbrv}
\bibliography{6.854-project}

\end{document}
